Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2020-08-27 Manuscript/HI_Model_ms.Rmd
updated manuscript files

 Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaii [@kim_population_2018]. For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016]. In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.

@@

Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well. However, Following @sleeter_effects_2019, we developed a set of separate scenarios designed to test the influence of a CFE on ecosystem carbon balance. We tested five $\beta$ levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments.

### *Scenario simulations*

@@

This study was funded by the Biological Sequestration Program of the U.S. Geological Survey. Thanks to Nicholas Koch of Forest Solutions, Inc. for information on eucalyptus harvesting and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software. Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S. Government.

@@

Tabular model output data are available from a USGS ScienceBase data repository: https://doi.org/10.5066/P9AWLFKZ. Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository: https://github.com/selmants/HI_Model.

>>>>

 Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawai`i [@kim_population_2018]. For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016]. In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.

@@

Increasing atmospheric CO~2~ concentrations stimulate leaf-level photosynthesis, potentially increasing NPP as well. However, the magnitude and persistence of this effect is highly uncertain, particularly across a range of climatic conditions and over long time spans. Following @sleeter_effects_2019, we developed a separate set of scenarios designed to test the sensitivity of LUCAS model ecosystem carbon balance projections to the influence of differing rates of a CO~2~ fertilization effect (CFE). We incorporated a NPP CFE multiplier representing the percent increase in NPP for every 100 ppm increase in atmospheric CO~2~ concentration under the high land use and high radiative forcing (RCP 8.5) scenario. We tested five CFE levels ranging from 5% to 15%, which is within the range of CFEs observed in free air CO~2~ enrichment (FACE) experiments. For all levels, we assumed CFEs reached saturation at an atmospheric CO~2~ concentration of 600 ppm, with no further increase in NPP despite a continued increase in CO~2~ concentration to 930 ppm by 2100. This 600ppm threshold generally coincides with the upper limit from FACE experiments and is reached by the year 2060 under RCP 8.5.

### *Scenario simulations and analysis*

@@

This study was funded by the Biological Sequestration Program of the U.S. Geological Survey. Thanks to Nicholas Koch of Forest Solutions, Inc. for information on eucalyptus harvesting and to Leonardo Frid and Colin Daniel of ApexRMS for assistance with SyncroSim software. Thanks also to Christian Giardina and Zhiliang Zhu for making this project possible. Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S. Government.

@@

Tabular model output data in machine readable format are available from a USGS ScienceBase data repository: https://doi.org/10.5066/P9AWLFKZ. Model input data and R code used to format input data, summarize output data, and compile this manuscript are available from a GitHub repository: https://github.com/selmants/HI_Model.
